An active sampling model of mental rotation

### 2014-01-27 man/cogsci-2014/cogsci-2014.tex
Fill in some citations

One of our most astonishing cognitive feats is the ability to
envision, manipulate, and plan with objects -- all without actually
perceiving them. This capacity for ``mental simulation'' has been
widely studied and has sparked intense debate about the underlying
representation of mental images \TODO{cite}. Despite this vast
literature, there have been surprisingly few computational models of
mental imagery.

@@

@@

evidence that people *don't* actually rotate until congruent

@@

In the mental rotation task, people are presented with pairs of
two-dimensional shapes similar to those used by \cite{Cooper:1975wp}
(e.g., Figure \ref{fig:stimuli}) and must determine whether the two
images depict the same shape, or mirror-image shapes.

@@

An alternative to the previous models is not to perform a search for
the MLE, but to approximate the integrals in Eq.~\ref{eq:lh-ratio} by
picking locations to sample which will most improve accuracy of the
estimate. Rather than a point estimate of the ratio, or even a simple
Monte Carlo estimate, we can infer a distribution by placing priors
over the similarity functions themselves:
\begin{equation}
\ell = \frac{\int \left[\int S(\Xb, f(\Xa, \theta, \hi))p(\theta)\dif\theta\right] p(S)\dif S}{\int \left[\int S(\Xb, f(\Xa, \theta, \hf))p(\theta)\dif\theta\right] p(S)\dif S}
\end{equation}
This is a technique known in the machine-learning literature as
\textit{Bayesian Quadrature} (BQ)
\cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm}. Specifically, we
use Gaussian Process (GP) priors \cite{Osborne:2012tm} \TODO{probably
  need to describe better what a GP is}. Because our domain is on a
circle, we utilize a periodic kernel for the covariance
\cite{Rasmussen:2006vz}, i.e. $k(\theta,
\theta^\prime)=h^2\exp[-\frac{2}{w^2}\sin^2(\frac{1}{2}(\theta-\theta^\prime))]$.

@@

Now that we have a distribution over $Z_h$, we can additionally
compute which points which we would expect to strengthen our estimate
(i.e., decrease the variance). From \cite{Osborne:2012tm}, the
expected variance of $Z_h$ if we sampled a new observation at
$\theta_a$ is:
\begin{align}
\mathbb{E}&[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) + \mathbb{E}[Z_h] - \\
&\int \mathbb{E}[Z_h|\theta_{a}]^2 \mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a, \theta_a))\dif\log S_h(\theta_a)\nonumber
\label{ref:expected-variance}
\end{align}

Revisiting our likelihood ratio, we now have:
\begin{equation}
p(\ell)\approx\frac{\mathcal{N}(\mathbb{E}[Z_0], \mathbb{V}(Z_0))}{\mathcal{N}(\mathbb{E}[Z_1], \mathbb{V}(Z_1))}
\end{equation}
\TODO{figure out what to do here.}

@@

We replicated the previous monotonic response time result

Also replicated the constant accuracy for flipped stimuli, but not for
same stimuli, from \TODO{cite cooper}. Why is this? Maybe the strategy
is ``they are different until proven the same''.

@@

>>>>

One of our most astonishing cognitive feats is the ability to
envision, manipulate, and plan with objects -- all without actually
perceiving them. This capacity for ``mental simulation'' has been
widely studied and has sparked intense debate about the underlying
representation of mental images \cite{Kosslyn:2009tj,
  Pylyshyn:2002vk}. Despite this vast literature, there have been
surprisingly few computational models of mental imagery.

@@

previous models of mental imagery: \cite{Kosslyn:1977tv, Anderson1978,
  Funt:1983wn, Glasgow:1992tj}

@@

evidence that people *don't* actually rotate until congruent,
e.g. rotating away from near-perfect matches \cite{Gardony:2013gn}

@@

In the mental rotation task, people are presented with pairs of
two-dimensional shapes similar to those used by \citeA{Cooper:1975wp};
an example stimulus is shown in Fig.~\ref{fig:stimuli}. The goal is to
determine whether the two images depict the same shape, or
mirror-image (flipped) shapes.

@@

An alternative to the previous models is not to perform a search for
the MLE, but to approximate the integrals in Eq.~\ref{eq:lh-ratio} by
picking locations to sample which will most improve accuracy of the
estimate. Rather than a point estimate of the ratio, or even a simple
Monte Carlo estimate, we can infer a distribution by placing priors
over the similarity functions themselves:
\begin{equation}
\ell = \frac{\int \left[\int S(\Xb, f(\Xa, \theta, \hi))p(\theta)\dif\theta\right] p(S)\dif S}{\int \left[\int S(\Xb, f(\Xa, \theta, \hf))p(\theta)\dif\theta\right] p(S)\dif S}
\end{equation}
This is a technique known in the machine-learning literature as
\textit{Bayesian Quadrature} (BQ)
\cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm}. Specifically, we
use Gaussian Process (GP) priors, as in \citeA{Osborne:2012tm}
\TODO{probably need to describe better what a GP is}. Because our
domain is on a circle, we utilize a periodic kernel for the covariance
\cite{Rasmussen:2006vz}, i.e. $k(\theta,
\theta^\prime)=h^2\exp[-\frac{2}{w^2}\sin^2(\frac{1}{2}(\theta-\theta^\prime))]$.

@@

Now that we have a distribution over $Z_h$, we can additionally
compute which points which we would expect to strengthen our estimate
(i.e., decrease the variance). From \citeA{Osborne:2012tm}, the
expected variance of $Z_h$ if we sampled a new observation at
$\theta_a$ is:
\begin{align}
\mathbb{E}&[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) + \mathbb{E}[Z_h] - \\
&\int \mathbb{E}[Z_h|\theta_{a}]^2 \mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a, \theta_a))\dif\log S_h(\theta_a).\nonumber
\label{ref:expected-variance}
\end{align}

Revisiting our likelihood ratio, we now have:
\begin{equation}
p(\ell)\approx\frac{\mathcal{N}(\mathbb{E}[Z_0], \mathbb{V}(Z_0))}{\mathcal{N}(\mathbb{E}[Z_1], \mathbb{V}(Z_1))}.
\end{equation}
This distribution does not have a nice form. However, we really are
just interested in whether $Z_0>Z_1$ or $Z_1>Z_0$: thus, rather than
computing $p(\ell)$, we use $Z_D=Z_0-Z_1$. Assuming $Z_0$ and $Z_1$
are independent,\footnote{This assumption is probably not
  correct. \TODO{explanation?}} we have:
\begin{equation}
  p(Z_D)\propto\mathcal{N}(\mathbb{E}[Z_0] - \mathbb{E}[Z_1], \mathbb{V}(Z_0) + \mathbb{V}(Z_1)).
\end{equation}
We then sample new observations until we are at least 95\% confident
that $Z_D\neq 0$. In other words, when $p(Z_D<0)<0.025$, we accept
$h=0$, and when $p(Z_D<0)>0.975$, we accept $h=1$.

@@

We replicated the previous monotonic response time result, though not
the linear resulte. Actually kind of similar to \cite{Gardony:2013gn}
\TODO{elaborate}

Also replicated the constant accuracy for flipped stimuli, but not for
same stimuli, from \cite{Cooper:1975wp}. Why is this? Maybe the
strategy is ``they are different until proven the same''.

@@

Other problems: evidence that rotations aren't actually holistic
\cite{Yuille:1982tx}, eye movements \cite{Just1976}
